\documentclass[12pt]{article}
\usepackage[latin1]{inputenc}
\usepackage{graphicx,subfigure}
\topmargin -1mm
\oddsidemargin -1mm
\evensidemargin -1mm
\textwidth 165mm
\textheight 220mm
\parskip 2mm
\parindent 3mm
%\pagestyle{empty}

\begin{document}

\begin{center}
{\Large \bf Getting started with OpenCaplhad}

Bo Sundman, \today

Opencalphad (OC) is a free software with GNU GPL license.  The last
version is available at the following web site:\\
http://www.opencalphad.org or the

A development version is available at the opencalpad repository at::\\
http://www.github.com/sundmanbo/opencalphad

You can contact me at: bo.sundman@gmail.com
\end{center}

{\large \bf Please be aware that this software is under development.
  Your feedback about problems and errors is important to make it
  better.  But if you have problems with the installation please
  contact a local expert.}

OC cannot replace your favorite thermodynamic software today or
tomorrow but the main advantage is that you have access to the source
code and can (with some efforts) add or fix things yourself that you
are missing in your favorite software.

There are several papers describing OC:

\begin{itemize}
\item {\bf OpenCalphad - a free thermodynamic software}, Bo Sundman,
  Ursula R Kattner, Mauro Palumbo and Suzana G Fries (Open Access)
  Integrating Materials and Manufacturing Innovation (2015) {\bf 4:1}
  DOI 10.1186/s40192-014-0029-1

\item {\bf The implementation of an algorithm to calculate
  thermodynamic equilibria for multi-component systems with non-ideal
  phases in a free software}, Bo Sundman, Xiao-Gang Lu and Hiroshi
  Ohtani, Computational Materials Science, {\bf 101} (2015) 127-137.

\item {\bf The OpenCalphad thermodynamic software interface}, Bo
  Sundman, Ursula R Kattner, Christophe Sigli, Matthias Stratmann,
  Romain Le Tellier, Mauro Palumbo and Suzana G Fries, Comp Mat Sci,
  {\bf 125} (2016) 188-196
\end{itemize}


There is also an extensive documentation provided with the software:

\begin{itemize}
\item Getting started with OpenCalphad (this document).
\item Install-OC, several different versions.
\begin{itemize}
\item Install-OC-Windows-MinGW on Windows using MinGW.
\item Install-OC-Windows-Cygwin on Windows using Cygwin (in French).
\item Install-OC-Linux on Linux using GNU.
\end{itemize}
\item news-OC4 about features and user interface.
\item UG, User Guide and manual.
\item OC-macros, describing the test macro files.
\item GTP, General Thermodynamic Pacakage describing the model package.
\item HMS, Hillert's Minimizer describing the equilibrium calculations.
\item SMP, StepMapPlot describing calculation and plotting of diagrams.
\item OCASI, OC Application Software Interface (draft).
\item Assessments using OC (draft).
\end{itemize}

\section{General information}

There is no automatic installation routine for OC, you must download
and compile the software yourself.  You may also have to install
Fortran compilers and the GNUPLOT software if you do not already have
them.  The OC development team cannot offer you any help for that,
please ask some local experts if you need help.

\begin{itemize}
\item The code is written in the new Fortran standard and requires a
  compiler like GNU Fortran 4.8 or similar.

\item The last release of the source code can be found as a zip file
  on http:://www.opencalphad.org.

  The last development version can be obtained as a zip file at the
  opencalphad repository on the http://www.github.com.  On that site
  there is also a wiki for information, discussion and questions.

\item See the special installation guides for advice on the
  installation.  If you have problems please contact a local expert.
  This is all free software and we are not experts on hardware.

\item If you have access to several CPUs you can test OC with
  parallelization using Open MP.

\item In the ``macro'' directory there are examples for a variety of
  calculations.

\item The TQ4lib directory has examples for using the software
  interface for Fortran and C/C++.

\item Contributions of new and improved source code are welcome.  You
  can do this by cloning the github repository.  Contact Bo Sundman if
  you want to know more.
\end{itemize}

\section{Some short hints to the user interface}

See also news-OC and UG.  The information in these (as well as this)
files are not always up-to-date.  You are welcome to inform me about
errors an inconsistencies.  If you do not like the current user
interface you are welcome to develop your own.

\subsection{Entering and manipulating thermodynamic data}

In this and the following sections you can find a brief and
unstructured summary how to use OC and of the commands that may be
useful.

\begin{itemize}
\item Thermodynamic data can be read from an unencrypted TDB file {\em
  read~tdb~``filename''} or entered interactively, see the {\em enter}
  commands.  You can specify which elements you want from the TDB file
  but not select the phases.  Phases you do not want you can suspend
  from the calculations with the command {\em
    set~status~phase~...~=~suspend}.

  If you already have data in the program when you read thermodynamic
  from a file you have to confirm that you want to overwrite the old
  data.  You can remove all the thermodynamic data in the program by
  the command {\em new~YES}.

\item There are some exception from how data are entered into OC
  compared to Thermo-Calc, most important perhaps the ``partitioning''
  of a phase into an ordered and disordered part.  OC has implemented
  this in a different way and at present it is not always possible for
  OC to read such partitioned data from a TDB file without editing the
  TDB file.

\item In this release you CANNOT save the thermodynamic data and
  results from a calculation on an unformatted file {\em
    save~UNFORM~``filename''} and read it back into the program with
  {\em read~UNFORM~``filename''}.  I will implement this and the
  DIRECT (random access) format in a future release.

\item You can save the thermodynamic data using the TDB format by a
  command\\ {\em list~data~TDB~filename}.  The TDB format used by OC
  can, with some exceptions, be read also by other thermodynamic
  software which accepts TDB format.

\item You can list the thermodynamic data on the screen with the {\em
  list~data} command.  You can list on a file by using the switch {\em
  list~/output=filename ...}.  If you want to write a TDB file use the
  command {\em list~data~TDB}.
  
\item You can enter thermodynamic data from a macro file as well as
  commands for calculations.  In order to document errors or problems
  please send a complete macro file with data and commands reproducing
  the error to Bo.Sundman@gmail.com

  You can generate a macro file interactively by setting a log file
  {\em set~log~``filename''} and then (after some editing) use the log
  file as a macro with the extension OCM.  The log file has the
  extension LOG.
\end{itemize}

\subsection{Setting conditions}

\begin{itemize}
\item To calculate you must set conditions on the external variables
  like temperature, T, pressure, P etc.  The number of conditions must
  be equal to the number of components plus 2 (Gibbs phase rule).

\item Setting conditions is similar to the Thermo-Calc software.  Each
  condition is set separately by the command is {\em set~cond~``state
    variable''~=~``value''}.  The safest set of conditions for
  calculating an equilibrium, i.e. which has most chance to converge,
  is to set values on T, P and N(element), i.e. the total amount of
  each element.  The table at the end gives a list of available state
  variable symbols and their meaning.

\item It is also possible to set conditions on chemical potentials,
  activities, enthalpies, volumes and that a phase is stable (fix).
  See the macro file examples how this is done.  Note that some
  commands are fragile and they may also change between releases of
  the OC software, depending on new ideas and suggestions by users.

\item The intention is that you should be able to combine any set of
  conditions to calculate the equilibrium, i.e. you should be able to
  combine conditions on mole fractions, mass fractions, fix phases,
  chemical potentials, enthalpies etc.  For conditions on mole
  fractions you can also enter expressions as conditions, for example
  to calculate the congruent melting of UO$_2$ you can set a condition
  x(liq,o)-x(c1\_mo2,o)=0 (see macro map4.OCM).

\item For users with FactSage background there is a command {\em
  set~input\_amount} which allows the user to set the overall
  composition by specifying the amount of moles of different species
  in the system.  The software converts this to {\em condtions} on the
  amounts of the components.

\item You can only use components to set conditions on the amounts,
  activities and chemical potential.   By default the components are
  the same as the elements but with the command {\em amend components}
  you can specify any orthogonal set of species as components.

  Beware that with other components than the elements your mole
  fractions and amounts of phases may be negative or larger than
  unity.

\end{itemize}

\subsection{Calculation}

\begin{itemize}
\item The command {\em calculate~equilibrium}, which can be
  abbreviated {\em c~e}, tries to calculate the equilibrium.  As the
  minimizer needs a guess of stable phases and their start
  constitution, the OC software tries first to invoke a global grid
  minimizer to find such a guess.  The grid minimizer requires that
  the conditions are $T, P$ and overall composition (N, B, x or w).

\item If you want to provide yourself a guess of the set of stable
  phases and constitutions you can use the {\em
    amend~const~phase~``name''~``amount''~``constitution'' $\cdots$}
  followed by the command {\em calculate~no\_global}, abbreviated {\em
    c~n}.

\item The grid minimizer that calculates start points for the general
  minimizer is powerful but a bit primitive.  If you have ideas how to
  improve it you are welcome to provide advice or code.  There is
  a command to have a denser grid (about 10 times more gridpoints) using
  the command {\em set~advanced~dense}

\item You should always check the results with you favorite
  thermodynamic software.  Please give documented feedback on any
  differences.

\item If the calculation does not converge directly try to use the
  command {\em ``c~n''} two or more times to continue to iterate from
  the set of phases you have.  The {\em ``c~e''} command will run the
  grid minimiser and, therefore, will normally give the same result
  each time.

\item You can also try to change the number of iterations and
  convergence criteria with the command {\em
    set~num~``max~iterations''~``convergce~criteria''}.  Default
  values of these are 500 and 10$^{-6}$.

\item If everything fails try to simplify the conditions for a first
  calculation.  The simplest are values of $T, P$ and $N(\rm
  component)$. When that has converged change the conditions one by
  one to those you are interested in and for each change calculate
  without the grid minimizer, i.e.  using {\em ``c~n''}.  Calculations
  at temperatures and compositions where the system is single phase
  have a higher chance of success.  The algorithm to change the set of
  stable phases is fragile and is still being fine-tuned.

\item Phases with miscibility gaps can be a problem.  The grid
  minimizer should detect such and automatically add new composition
  sets for phases that can be stable with two different compositions.
  Composition sets created automatically have the suffix AUTO.

\item The user can also add composition sets manually with the
  command\\ {\em amend~phase~``name''~comp\_set }.  You can add a
  prefix and suffix (max 4 letters) to the phase name (for example the
  cubic carbide treated as a composition set to the FCC phase can have
  the prefix MC so the full name is MC\_FCC.

  Composition sets have a number after the phase name preceded by the
  letter \#, like LIQUID\#2.  The user can use both pre- and suffixes
  or use the composition set number to specify the composition set.  A
  phase can have up to 9 composition sets.

\item When entering a composition set manually you are asked for a min
  or max values of each constituent.  After a calculation the software
  will try to match a calculated constitution of a phase which has
  several composition sets to the set with the closest default
  constitution.

\item To set the default constitution of the first composition set use
  {\em set~phase~``name''~default\_constitution~$\cdots$ }.

\item There are several options to calculate things:
  \begin{itemize} 
  \item {\em calculate~all} calculates all equilibria set by the {\rm
    set~range} command.  It is intended for equilibria with
    experiments and used in assessments but can be used to calculate a
    range of equilibria not suitable for the STEP command.
  \item {\em calculate global\_gridmin} will use the global
    gridminimizer and not the final iterative routine,
  \item {\em calculate~phase} to calculate properties for a specific phase.
    You must first specify the amount of the phase and if you want to
    use the current constitution.  Then you can select
    \begin{itemize}
    \item {\em only G} to get G and derivatives with respect to $T$.
    \item {\em G and dGdy} to get G and also first derivatives
      w.r.t. constituents.
    \item {\em all derivatives} to get G and all first and second
      derivatives w.r.t $T$, $P$ and constituents.
    \item {\em constitution adjustment} This will ask for a new
      composition (default is the current or the one specified) and
      miminize the Gibbs energy of the specified phase to have the
      equilibrium constitution (for example adjust the fraction of gas
      constituents or sublattice fractions).  It will return the value
      of G and all chemical potentials (divided by RT).
    \item {\em diffusion coeff} This will also ask for a new
      composition and return the values of the Darken stability matrix
      calculated as $\frac{\partial G_{\rm A}}{\partial N_{\rm B}}$
      and if the phase has any mobility data also their values.
    \end{itemize}
  \item {\em calculate~symbol} will calculate the value of one or more
    symbols.
  \item {\em calculate~tp\_funs} will calculate the value and the
    first and second derivatives with respect to $T$ and $P$ for all
    tp functions (used for the model parameters).
  \item {\em calculate~transition~``phase''~``condition''} calculates
    the value of the ``condition'' when the phase is just stable, for
    example the melting $T$ if the ``phase'' is LIQUID.
  \end{itemize}
\end{itemize}

\subsection{Setting the status of phases: entered, fixed, dormant, suspended}

To set the status of a phase to be fixed is treated as a condition.

\begin{itemize}
\item By default all phases have the status ENTERED.  You can suspend
  phases you do not want by {\em
    set~status~phase~``name''~$\cdots$~=~``suspend''}.

  If the phase should be stable you will calculate a metastable equilibrium.

\item To specify that a phase must be stable you can set its status to
  fix, the command is

  {\em set~status~phase~``name''~=~``fix''~''amount''}.

  If the amount is set to zero OC will calculate a point on the
  solubility line of the phase.  Note it may fail because there is no
  such point due to the other conditions.

\item The dormant status means the phase cannot be stable but its
  driving force is calculated.  If this is positive the phase wants to
  be stable and the calculated equilibrium is metastable.
\end{itemize}

\subsection{Listing many things}

Most listing are written on the screen but when output obviously
should be written on a file the program will ask for a file name.  If
you want the output from other listings to be written on a file you
can use the switch {\em /output=filename} or {\em /append=filename}
directly after the first command.  For example {\em list~/out=result~r,,,}.
The extension of the output file will be DAT.  The output file will be
reset to the screen after each command.

There is also a SAVE command mainly intended for saving on a file.
For example {\em save tdb filename} will write all the thermodynamic
data in the OC TDB format on the file with the extension TDB.  This
can be read back later as a TDB file.

\begin{itemize}
\item Listing results after a calculation is done by {\em list~result
  ``option''}.  This is the default list and can be abbreviated {\em
  l,,}.  This list will always include the current conditions and some
  global results and one line for each component.  The options are at
  present:
  \begin{itemize}
    \item 1 list of stable phases with amount and composition in mole
      fractions in value order.
    \item 2 list of stable phases with amount, composition and
      constitution in mole fractions in value order.
    \item 3 list of stable phases with amount and composition in mole
      fractions in alphabetical order.
    \item 4 list of stable phases with amount and composition in
      mass fractions in value order.
    \item 5 list of stable phases with amount and composition in
      mass fractions in alphabetical order.
    \item 6 list of stable phases with amount and composition in
      mass fractions and the constitution in value order.
    \item 7 list of all phases with amount and composition in
      mass fractions in value order.
    \item 8 list of all phases with amount and composition in
      mole fractions and constitution in alphabetical order.
    \item 9 list of all phases with amount and composition in
      mole fractions and constitution in value order.
  \end{itemize}
\item The {\em list~short~A} gives one line for each component,
  species and phase.  In this list you may note the use of the ``phase
  tuple'' to indicate composition sets.  A phase tuple has two values,
  one is the phase number, the other the composition set number.  The
  first composition set of a phase has the same tuple number as the
  phase number.  As composition sets can be added and deleted the
  tuple number of a higher composition set may change.
\item The {\em list~short~P} gives one line for each phase, separated
  in stable, entered and dormant, listed in order of increasing
  instability (driving force).
\item The {\em list~bibliography} list all or a single bibliographic
  reference.  Bibliographic references are entered when entering
  parameters.  The {\em amend~biblio} allows interacting amending of
  the bibliographic item.
\item The {\em list~data~screen} lists all thermodynamic data with
  bibliographic references.  Note this output is not the TDB format.
\item The {\em list~equilibria} list all entered equilibria (not
  any results).  By default there is only one equilibrium but during step
  and map each node point is entered as an equilibria.  The user can
  also enter equilibria manually.
\item The {\em list~model\_param\_id} list all symbols for which one
  can enter parameters.  By default the G is Gibbs energy, TC the
  combined Curie and Neel temperature etc.
\item The {\em list~phase} has several options to list various data
  for a phase.
\item The {\em list~state\_variables} lists individual values of state
  variables.  The wildcard character ``*'' can be used for ``all''.
  Terminate with an empty line.
\item The {\em list~symbols} list the symbols and their expressions.
  There two predefined symbols, R and RT.  The user can enter any
  number of additional symbols as functions of state variables or
  other symbols.  To calculate (and list) the value of a symbol use
  {\em c~sym}.
\item The {\em list~tp\_symb} lists the functions entered for the
  model parameters.  To calculate their current values use {\em c~tp}.
\end{itemize}

\subsection{Calculating and plotting diagrams}

This part is unfinished and very rudimentary.

\begin{itemize}
\item You can calculate property diagram (one independent axis
  variable) or phase diagrams (two or more independent axis variables).
  The command to set an independent axis is 

  {\em set~axis~1~``state~variable''~``min value''~``max
    value''~``increment''}.  The state variable used must be a
  condition.  Axis 1 and 2 are implemented.

\item With one axis the {\em step} command will step from the current
  equilibrium to the max and min axis values.

\item With two axis the command {\em map} will follow ``zero phase
  fraction'' (ZPF) lines within the limits of the axis by replacing
  one of the axis condition with a phase fix with zero amount.  In
  this way phase diagrams for any number of components can be
  calculated.  While following such a line other phases may become
  stable or disappear and this generates a node point where several
  lines meet.

\item Plotting of diagrams relies on GNUPLOT, a free software you can
  install from the web.  The command {\em plot} asks for two axis
  variables and generates pairs of values from the results calculated
  by step or map commands.

  There are several options to the plot command that are demonstrated by
  the macro files.

\item Macro files are easily created by using a log file while running
  OC interactively.  On the log file all input (as well as default
  values) are written.  Then you can edit this and finish it with the
  {\em set interactive} command to give command back to the keyboard.

  When you report problems or errors always sent a macro file that
  reproduces the error.  Please try to find the simplest case that
  contains the error.
\end{itemize}

\section{A summary of state variables.}

The state variables are very important for setting conditions and
listings.  Those recognized by OC are given in
table~\ref{tab:statevar}.

\begin{table}[!h]
\caption{A very preliminary table with the state variables and their
  internal representation.  Some model parameter properties are also
  included.  The final ``z'' in Sz etc is the optional normallizing
  letter M, W, V or F}\label{tab:statevar}
\begin{tabular}{|llccll|}\hline
Symbol & Id & \multicolumn{2}{c}{Index} & Normalizing & Meaning\\
       &    & 1 & 2                     &  suffix     & \\\hline
\multicolumn{6}{|c|}{Intensive properties}\\\hline
T      & 1  & -         & -    & - & Temperature\\
P      & 2  & -         & -    & - & Pressure\\
MU     & 3  & component & -/phase  & - & Chemical potential\\
AC     & 4  & component & -/phase  & - & Activity\\
LNAC   & 5  & component & -/phase  & - & LN(activity)\\\hline
\multicolumn{6}{|c|}{Extensive and normallized properties}\\\hline
U      & 10 & -/phase\#set & - & - & Internal energy for system\\
UM     & 11 & -/phase\#set & - & M & Internal energy per mole\\
UW     & 12 & -/phase\#set & - & W & Internal energy per mass\\
UV     & 13 & -/phase\#set & - & V & Internal energy per m$^3$\\
UF     & 14 & phase\#set   & - & F & Internal energy per formula unit\\
Sz     & 2z & -/phase\#set & - & - & entropy\\
Vz     & 3z & -/phase\#set & - & - & volume\\
Hz     & 4z & -/phase\#set & - & - & enthalpy\\
Az     & 5z & -/phase\#set & - & - & Helmholtz energy\\
Gz     & 6z & -/phase\#set & - & - & Gibbs energy\\
NPz    & 7z &  phase\#set & - & - & Moles of phase\\
BPz    & 8z & phase\#set & - & - & Mass of phase\\
Qz     & 9z & phase\#set & - & -  & Stability of phase\\
DGz    & 10z & phase\#set & - & -  & Driving force of phase\\
Nz     & 11z & -/phase\#set/comp & -/comp & -  & Moles of component\\
X      & 111 & phase\#set/comp & -/comp & 0  & Mole fraction\\
X\%    & 111 & phase\#set/comp & -/comp & 100 & Mole per cent\\
Bz     & 12z & -/phase\#set/comp & -/comp & -  & Mass of component\\
W      & 122 & phase\#set/comp & -/comp & 0 & Mass fraction\\
W\%    & 122 & phase\#set/comp & -/comp & 100 & Mass per cent\\
Y      & 130 & phase\#set & const\#subl & -& Constituent fraction\\\hline
\multicolumn{6}{|c|}{Some model parameter identifiers}\\\hline
TC     & - & phase\#set & - & - & Curie temperature\\
BMAG   & - & phase\#set & - & - & Aver. Bohr magneton number\\
MQ\&X  & - & phase\#set & constituent X & - & Mobility\\
THET   & - & phase\#set & - & - & Debye temperature\\\hline
\end{tabular}
\end{table}

The state variables in the user interface have their common symbols,
$T$ for temperature, $P$ for pressure, $N$ for the total amount of
moles, ``$N$(element)'' for the amount of moles of a component,
``$X$(element)'' for the mole fraction ``$MU$(element)'' for the
chemical potential, ``$AC$(element)'' for the activity.  The symbol
$B$ is used for the total mass (copied from the Thermo-Calc software),
``$B$(element)'' for the mass of an element and ``$W$(element)'' for
the mass fraction.  There are many more state variables like $H$, $G$
etc, see the table, but not all of them can be used as conditions.

\section{Phase status}

You can specify that a phases should be stable by the command {\em set
  status phase $\cdots$}.  For example to calculate the melting point
of an alloy after specifying the composition and making a calculation
at fixed T and P, you can give the commands {\em set cond T=none; set
  status liquid=fix 0; c n}.  (The commands must be given on separate
lines).  You can also use the command {\em calculate transition
  $\cdots$}.

\section{Manipulating the source code}

The OC software is provided with a GNU license which means that you
have the source code and can use it and modify it as you wish as long
as you do not try to make money of it.  If you want to include the OC
software in a commercial program you must contact the copyright
owners.

There is a fairly extensive documentation of the source code in the
directory ``documentation'' and if you look at the code itself there
are some comments there too.  I have spent a lot of effort to make the
datastructures general and flexible to handle multicomponent and
multiphase systems.  But there was quite a lot of redundancy
introduced during the development that eventually will be removed.

The set of subroutines is less structured and one problem has been
that this code was my first attempt to use the new Fortran standard so
there are probably many things that can be made simpler.  You are
welcome to point out where this can be done.  The release of version 3
will be in January 2016 and the work on an improved version will
continue.

As I have understood the data structures (TYPE) in the new Fortran
standard is more or less identical to those used in C++ so it should
not be too difficult to combine code written in these languages.

\section{Application software library, OCACI}

There is a software interface following the TQ standard to make it
possible to use OC from application programs.  The subroutines in this
interface can be found in the directory ``TQ4lib'' together with a few
test program.  Read the separate readme files in these directories to
compile and run these.

There is a Fortran version of the TQ library and also an isoC version
that can be called from C++/C routines.  The isoC binding makes it
possible to access data directly from the datastructures defined
inside the OC software.  There are some test programs for this also.

Use only subroutines in the TQ library to access the OC software, do
not call directly subroutines inside the OC code as they may not be
available or have different functionality in a future release.  If
you miss some routines please sent a message or discuss this on the
github wiki.

\section{Parallelization}

The code has been developed with parallelization in mind and I have
now started to test this code and surprisingly it seems to work quite
well.  There is a macro parallel2.OCM showing a calculation of 400
different equilibria in a 6 component alloy.  They need to be
initialized which makes me a bit suspicious but testing changing some
and calculate again seems to work.  I have tested on a PC with 4 CPU
and the clock time is about 3.5 times faster compared to calculate
them sequentially.

Parallization can be implemented in several levels in OC, for example
the grid minimizer, but the ability to calculate many separate
equilibria in parallel is a key feature to implement full
thermodynamic calculations in simulations.

\bigskip

{\large \bf Have fun and help make OC useful!}

\end{document}

